---
title: "3 - Data Visualization"
output: html_notebook
---

Load the `tidyverse` library:

```{r}
library(tidyverse)
```

The `mpg` dataset contains observations from various models of cars:

```{r}
mpg
```


We can plot the highway mpg vs. engine size:

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

3.2.4 Exercises

1. Run `ggplot(data = mpg)`. What do you see?

```{r}
ggplot(data = mpg)
```

We see a blank ggplot canvas since we have not added any geometric objects.

2. How many rows are in `mpg`? How many columns?

```{r}
dim(mpg)
```

In the `mpg` dataset, the number of rows is 234, and the number of columns is 11.

3. What does the `drv` variable describe? Read the help for `?mpg` to find out.

```{r}
?mpg
```

The `drv` variable describes whether the vechile is front-wheel drive (f), read-wheel drive (r), or four-wheel drive (4).

4. Make a scatterplot of `hwy` vs `cyl`.

```{r}
ggplot(data = mpg) +
  geom_point(aes(x = cyl, y = hwy))
```

5. What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?

```{r}
ggplot(data = mpg) +
  geom_point(aes(x = cyl, y = class))
```

This plot is since multiple classes have multiple numbers of cylinders.

We can plot the highway mpg vs. engine size and display the color of the points based on the vehicle class:

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

3.3.1 Exercises

1. What’s gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

The `color` option should be outside of the aesthetic settings.

2. Which variables in `mpg` are categorical? Which variables are continuous? (Hint: type `?mpg` to read the documentation for the dataset). How can you see this information when you run `mpg`?

```{r}
mpg
```

In the `mpg` dataset, categorical values such as `model`, `displacement`, and `cyl` are prescribed, while continuous variables such as `cty` and `hwy` are measured. You can view the type information for the variables in the dataframe as a guide.
